Uncovering underlying high-level musical content in the time domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn deep audio embeddings with applications to boundary detection tasks

### 2023-07-06 acknowledgements/acknowledgements.tex
Updates from Overleaf

\begin{itemize}
\item Carl Thomé and Carlos Lordeloª, whose expertise in MIR, ML, and DSP has been pivotal to the progress and success of my thesis. Their invaluable guidance and unwavering encouragement have empowered me to overcome the challenges encountered throughout this journey. Their dedication and commitment to mentorship are truly inspiring, and their influence has left an indelible mark on my academic path. The opportunity to work under their mentorship has been a privilege and an honor. Their profound impact is beyond what words can encapsulate.

@@

@@

\vspace*{3mm}
\item My colleague at My Sheet Music Transcriptions, Oriol López, for his exceptional support and flexibility throughout my thesis. His understanding and willingness to accommodate my academic commitments effectively allowed me to balance my work and research responsibilities.

\vspace*{3mm}
Thank you all for being excellent advisors and playing a crucial role in completing my thesis. I sincerely appreciate everything you have done for me and will always cherish the knowledge and experience I have gained under your guidance.

>>>>

\begin{itemize}
\item Carl Thomé and Carlos Lordelo, whose expertise in MIR, ML, and DSP has been pivotal to the progress and success of my thesis. Their invaluable guidance and unwavering encouragement have empowered me to overcome the challenges encountered throughout this journey. Their dedication and commitment to mentorship are truly inspiring, and their influence has left an indelible mark on my academic path. The opportunity to work under their mentorship has been a privilege and an honor. Their profound impact is beyond what words can encapsulate.

@@

\vspace*{3mm}
\item Oriol Nieto and J. Smith

@@

\vspace*{3mm}
\item My colleague at My Sheet Music Transcriptions, Oriol López, for his exceptional support and flexibility throughout my thesis. His understanding effectively allowed me to balance my work and research responsibilities.

\vspace*{3mm}

Thank you all for playing a crucial role in completing my thesis. I sincerely appreciate everything you have done for me and will always cherish the knowledge and experience I have gained under your guidance.

### 2023-07-06 introduction/Subsections/Introduction.tex
Updates from Overleaf

Beyond applications in music structure analysis, the evolution of neural networks has empowered the creation of condensed numerical representations encapsulating crucial musical traits, technically known as representation (or feature) learning. This evolution has simplified the computational process for various task-specific elements, including but not limited to:

\begin{itemize}
\item Creating an audio embedding method that excels across numerous applications without the necessity for fine-tuning \cite{Turian2022HEAR:Representations}.
\item Boosting the classification of environmental sounds \cite{CramerLOOKEMBEDDINGS}.
\item Enhancing vocal-centric music tasks through cross-domain audio embeddings \cite{Kim2021LearningLoss}.
\item Integrating task-specific and pre-trained features to optimize audio classification \cite{Hung2022Feature-informedClassification}.
\item Developing a music similarity search engine for video production \cite{epidemic}.
\item Improving Music Emotion Recognition (MER) performance \cite{KohComparisonRecognition}.
\item Evaluating the effectiveness of speaker recognition via pre-trained model embeddings \cite{lightweight}.
\item Embedding songs for artist identification to facilitate similarity comparisons \cite{contentmusicsimtriplet2020}.
\item Addressing cross-modal text-to-music retrieval issues \cite{WonEmotionStories}.
\item Enabling automated music rearrangement \cite{Stoller2018IntuitiveTransitions, Plachouras2023MusicSegmentation}.
\end{itemize}

Furthermore, deep audio embeddings offer the benefit of transferability, establishing a foundation for multiple tasks. Compared to training a model from scratch, this approach saves computational resources and time \cite{transferMIR2013, CifkaDeepTransfer}.

>>>>

Beyond applications in music structure analysis, the evolution of neural networks has empowered the creation of latent representation encapsulating crucial musical traits, technically known as representation (or feature) learning. This evolution has simplified the computational process for various task-specific elements, including but not limited to:

\begin{itemize}
\item Creating an audio embedding that excels across numerous applications without the necessity for fine-tuning \cite{Turian2022HEAR:Representations}.
\item Boosting the classification of environmental sounds \cite{CramerLOOKEMBEDDINGS}.
\item Enhancing vocal-centric music tasks through cross-domain audio embeddings \cite{Kim2021LearningLoss}.
\item Integrating task-specific and pre-trained features to optimize audio classification \cite{Hung2022Feature-informedClassification}.
\item Developing a music similarity search engine for video production \cite{epidemic}.
\item Improving Music Emotion Recognition (MER) performance \cite{KohComparisonRecognition}.
\item Evaluating the effectiveness of speaker recognition via pre-trained model embeddings \cite{lightweight}.
\item Embedding songs for artist identification to facilitate similarity comparisons \cite{contentmusicsimtriplet2020}.
\item Addressing cross-modal text-to-music retrieval issues \cite{WonEmotionStories}.
\item Enabling automated music rearrangement \cite{Stoller2018IntuitiveTransitions, Plachouras2023MusicSegmentation}.
\end{itemize}

Furthermore, deep audio embeddings offer the benefit of transferability, establishing a foundation for multiple tasks. Compared to training a model from scratch, this approach saves computational resources and time \cite{transferMIR2013, CifkaDeepTransfer, Ding2023AudioClassification}.

### 2023-07-06 introduction/Subsections/Objectives.tex
Updates from Overleaf

Can deep audio embeddings be learned from reordering, scrambling, and augmenting via audio effects sequences of musical information to improve unsupervised music boundary detection?

This study explores the high-level tonal structure of music by leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn and analyze music embeddings that can be applied to subsequent MIR tasks, with a particular focus on music boundary detection.

>>>>

Can deep audio embeddings be learned from reordering, scrambling, and augmenting sequences of musical information to improve unsupervised music boundary detection?

### 2023-07-06 introduction/Subsections/Previous work.tex
Updates from Overleaf

Related research \cite{deepfeaturesegment, SalamonDeepSegmentation} investigates the use of audio embeddings derived through unsupervised methods to enhance the performance of music segmentation algorithms. Both studies leverage the power of deep learning and data-driven feature learning, presenting advancements over traditional manually-engineered methods.

@@

The results prove that unsupervised deep learning techniques that derive audio embeddings offer a robust and more efficient alternative to traditional manual feature engineering methods in music segmentation.

>>>>

Related research investigates using audio embeddings derived through unsupervised methods to enhance the performance of music segmentation algorithms \cite{deepfeaturesegment, SalamonDeepSegmentation}. Both studies leverage the power of deep learning and data-driven feature learning, presenting advancements over traditional manually-engineered methods.

@@

The results prove unsupervised deep learning techniques that derive audio embeddings offer a robust and more efficient alternative to traditional manual feature engineering methods in music segmentation.

### 2023-07-06 methods/subsections/implementation.tex
Updates from Overleaf

Our implementation offers a slight modification to the \textit{contrastive learning of musical representations} (CLMR), a method that learns valuable, discriminative music representations without explicit labels by contrasting positive augmentations of a musical piece against negative ones \cite{CLMR2021}. CLMR falls under a branch of ML known as self-supervised learning (SSL) \cite{Balestriero2023ALearning}, where models learn autonomously from unlabeled data to create their own supervisory signals \cite{audioselfsupsurvey}. This method resembles how humans learn from observations and interactions, transforming unsupervised problems into supervised ones by auto-generating labels. SSL benefits include reduced dependence on labeled data, contributing to more robust and generalizable data representations.

@@

\begin{table}[ht]
\centering
\small
\begin{tabularx}{\textwidth}{>{\hsize=1.6\hsize}X>{\hsize=0.6\hsize}X>{\hsize=0.6\hsize}X}
\toprule
\thead{\textbf{Layer Type}} & \thead{\textbf{In Channels}} & \thead{\textbf{Out Channels}} \\
\midrule
Conv + ReLU & 1 & 128 \\
\addlinespace
\multicolumn{3}{c} \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 128 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 128 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 512 \\
\addlinespace
Conv + BatchNorm + ReLU + AvgPool & 512 & 512 \\
\addlinespace
Fully connected layer & 512 & 128 \\
\bottomrule
\end{tabularx}
\caption{Layer specifications for SampleCNN model}
\label{tab:samplecnntable}
\end{table}

>>>>

Our implementation offers a slight modification to the \textit{contrastive learning of musical representations} (CLMR) method, which learns valuable, discriminative music representations without explicit labels by contrasting positive augmentations of a musical piece against negative ones \cite{CLMR2021}. CLMR falls under a branch of ML known as self-supervised learning (SSL) \cite{Balestriero2023ALearning}, where models learn autonomously from unlabeled data to create their own supervisory signals \cite{audioselfsupsurvey}. This method resembles how humans learn from observations and interactions, transforming unsupervised problems into supervised ones by auto-generating labels. SSL benefits include reduced dependence on labeled data, contributing to more robust and generalizable data representations.

@@

\begin{table}[ht]
\centering
\small
\begin{tabularx}{\textwidth}{
  >{\centering\arraybackslash\hsize=1.6\hsize}X
  >{\centering\arraybackslash\hsize=0.6\hsize}X
  >{\centering\arraybackslash\hsize=0.6\hsize}X}
\toprule
\thead{\textbf{Layer Type}} & \thead{\textbf{In Channels}} & \thead{\textbf{Out Channels}} \\
\midrule
Conv + ReLU & 1 & 128 \\
\addlinespace
\multirow{9}{*}{Conv + BatchNorm + ReLU + MaxPool} & 128 & 128 \\
\addlinespace
 & 128 & 128 \\
\addlinespace
 & 128 & 256 \\
\addlinespace
 & 256 & 256 \\
\addlinespace
 & 256 & 256 \\
\addlinespace
 & 256 & 256 \\
\addlinespace
 & 256 & 256 \\
\addlinespace
 & 256 & 256 \\
\addlinespace
 & 256 & 512 \\
\addlinespace
Conv + BatchNorm + ReLU + AvgPool & 512 & 512 \\
\addlinespace
Fully connected layer & 512 & 128 \\
\bottomrule
\end{tabularx}
\caption{Layer specifications for the SampleCNN model}
\label{tab:samplecnntable}
\end{table}

### 2023-07-06 preface/preface.tex
Updates from Overleaf

This work humbly aims to help untangle the complexities of music and offer a more musically-driven perspective to MIR. We aim to contribute to a deeper understanding of the intricate relationships within musical data by bridging the gap between music theory and computational analysis.

@@

This work is intended for any curious human being, regardless of their background in music or computer science. I hope it will inspire others to explore the fascinating intersection of these two disciplines.

>>>>

This work humbly aims to help untangle the complexities of music and offer a more musically-driven perspective to MIR. It aims to contribute to a deeper understanding of the intricate relationships within musical data by bridging the gap between music theory and computational analysis.

@@

This work is intended for any curious person, regardless of their academic background. I hope it will inspire others to explore the fascinating intersection of these two disciplines.

### 2023-07-06 results/subsections/tables.tex
Updates from Overleaf

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\centering\arraybackslash}p{4.5cm}XXXX}
\toprule
\thead{\textbf{Authors [Ref], Year}} & \thead{\textbf{Input}} & \thead{\textbf{Method}} & \thead{\textbf{F-measure}} \\
\midrule
Turnbull et al. \cite{Turnbull2007ABOOSTING}, 2007 & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
Kaiser et al. \cite{27}, 2012 & SSM & Novelty measure  & 0.286 \\
\addlinespace
McFee \& Ellis \cite{20}, 2013 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
Nieto \& Bello \cite{28}, 2014 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2015 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
Nieto \cite{30}, 2016 & CQT Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2017 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
Sargent et al. \cite{34}, 2011 & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
Ullrich et. al \cite{22}, 2014 & MLS & CNN  & 0.465 \\
\addlinespace
Grill \& Schlüter \cite{4}, 2015 & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
Grill \& Schlüter \cite{Grill2015MusicAnnotations}, 2015 & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
Hadria \& Peeterss \cite{35}, 2017 & MLS, SSLMs & CNN  & 0.291 \\
\addlinespace
\cite{deepfeaturesegment}, 2019 & CQT & CNN  & \textbf{0.535} \\
\addlinespace
Ours, 2023 & Raw waveforms & CNN  & 0.288 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State-of-the-art table.]{Previous studies' boundary detection f-measure results using unsupervised methods for a 0.5s time-window tolerance. Only the top-performing algorithm for each year on the SALAMI dataset is displayed. This table has been extended from \cite{Hernandez-Olivan2021MusicFeatures}.}
\label{tab:comparison_table}
\end{table}

>>>>

\begin{table}[ht]
\begin{threeparttable}
\centering
\small
\begin{tabularx}{\textwidth}{
  >{\centering\arraybackslash}p{4.5cm}
  >{\centering\arraybackslash}X
  >{\centering\arraybackslash}X
  >{\centering\arraybackslash}X}
\toprule
\thead{\textbf{Authors [Ref], Year}} & \thead{\textbf{Input}\tnote{1}} & \thead{\textbf{Method}} & \thead{\textbf{F-measure}} \\
\midrule
Turnbull et al. \cite{Turnbull2007ABOOSTING}, 2007 & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
Kaiser et al. \cite{27}, 2012 & SSM & Novelty measure  & 0.286 \\
\addlinespace
McFee \& Ellis \cite{20}, 2013 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
Nieto \& Bello \cite{28}, 2014 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2015 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
Nieto \cite{30}, 2016 & CQT Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2017 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
Sargent et al. \cite{34}, 2011 & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
Ullrich et. al \cite{22}, 2014 & MLS & CNN  & 0.465 \\
\addlinespace
Grill \& Schlüter \cite{4}, 2015 & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
Grill \& Schlüter \cite{Grill2015MusicAnnotations}, 2015 & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
Hadria \& Peeterss \cite{35}, 2017 & MLS, SSLMs & CNN  & 0.291 \\
\addlinespace
McCallum \cite{deepfeaturesegment}, 2019 & CQT & CNN  & \textbf{0.535} \\
\addlinespace
Ours, 2023 & Raw waveforms & CNN  & 0.288 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State-of-the-art table.]{Previous studies' boundary detection f-measure results using unsupervised methods for a 0.5s time-window tolerance. Only the top-performing algorithm for each year on the SALAMI dataset is displayed. This table has been extended from \cite{Hernandez-Olivan2021MusicFeatures}.}
\label{tab:comparison_table}
\begin{tablenotes}\footnotesize
\item[1] \textbf{Legend:} SSM: Self-Similarity Matrix, MLS: Mel Spectrogram, MFCC: Mel-Frequency Cepstral Coefficient, CQT: Constant Q Transform, PCP: Pitch Class Profile, SSLM: Self-Similarity Lag Matrix.
\end{tablenotes}
\end{threeparttable}
\end{table}
